---
title: "NC Analysis of Mock U of Michigan Health System Data"
author: "Kendrick Li"
date: '2022-11-11'
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(dplyr)
library(MASS)
library(lubridate)
library(numDeriv)
source("nc_functions.R")
```

In this document, we illustrate the proposed NC estimator using a mock University of Michigan Health System data to study SARS-Cov-2 vaccine effectiveness against symptomatic infection.

## Setup and load the dataset
```{r cars}

sessionInfo()

tnd_data <- read.csv(file = "mock_tnd_data.csv",
                     header = TRUE,
                     stringsAsFactors = FALSE)
with(tnd_data, table(Vaxname, COVID))
```

## Data Preprocessing
1. Categorize age into age groups (younder than 18, 18 to 60, older than 60);
2. Create dummy variables for Charlson Score (0, 1, 2, larger than 3, missing);
3. Create dummy variables for calendar months of the latest COVID test;
4. Create an indicator for Caucasian subjects;
5. Create the negative control outcome (NCO) as a binary indicator of the presence of at least one of the following conditions: arm/leg cellulitis, eye/ear disorder,  gastro-esophageal disease, atopic dermatitis, injuries, and general adult examination visits.
```{r pressure, echo=TRUE}
tnd_data_2 <- tnd_data %>%
  mutate(age_le_18 = as.numeric(Age < 18),
         age_geq_18_le_60 = as.numeric(Age >= 18 & Age < 60),
         age_geq_60 = as.numeric(Age >= 60),
         wscore_0 = as.numeric(!is.na(wscore) & wscore == 0),
         wscore_1 = as.numeric(!is.na(wscore) & wscore == 1),
         wscore_2 = as.numeric(!is.na(wscore) & wscore == 2),
         wscore_geq_3 = as.numeric(!is.na(wscore) & wscore >= 3),
         wscore_na = as.numeric(is.na(wscore)),
         april = as.numeric(months == 4),
         may = as.numeric(months == 5),
         june = as.numeric(months == 6),
         july = as.numeric(months == 7),
         aug = as.numeric(months == 8),
         sep = as.numeric(months == 9),
         oct = as.numeric(months == 10),
         nov_dec = as.numeric(months >= 11),
         caucasian = as.numeric(Race == "Caucasian"),
         nco = as.numeric(EE + CLT + gfd + dermatitis + injuries + GE_NCO > 0))
```

## Detecting the confounding bias
1. Adjusted logistic regression of the NCO on vaccination and NCE 
```{r nco_reg, echo = TRUE}
glm_NCO <- glm(nco ~ vaccinated + IMM + age_geq_18_le_60 + age_geq_60 + Gender + caucasian + wscore_geq_3 + april + may + june + july + aug + sep + oct,
               family = binomial,
               data = tnd_data_2)
summary(glm_NCO)
```
The p-values for the coefficients of SARS-Cov-2 vaccination and NCE (prior immunization) are <2e-16 and 2.09e-14, respectively, indicating the presence of confounding bias.

2. Adjusted logistic regression of COVID-19 on vaccination and NCE 
```{r nce_reg, echo = TRUE}
glm_NCE <- glm(COVID ~ vaccinated + IMM + age_geq_18_le_60 + age_geq_60 + Gender + caucasian + wscore_geq_3 + april + may + june + july + aug + sep + oct,
               family = binomial,
               data = tnd_data_2)
summary(glm_NCE)
```
The p-value for the coefficient NCE (prior immunization) is <2e-16, also indicating the presence of confounding bias.

## NC estimator of vaccine effectiveness
1. Pfizer-BioNTech vaccine
```{r pfizer_nc}
pfizer_data <- tnd_data_2 %>% filter(Vaxname %in% c("Pfizer", "none"))

pfizer_iptw <- tnd_iptw(data = list(A = pfizer_data$vaccinated,
                                    Y = pfizer_data$COVID,
                                    X = pfizer_data[, c("age_geq_18_le_60", "age_geq_60", "Gender", "caucasian", "wscore_geq_3", "april", "may", "june", "july", "aug", "sep", "oct")]))

pfizer_nc <- nc_covid_ve(data = pfizer_data,
                         nce = "IMM",
                         nco = "nco",
                         x = c("age_geq_18_le_60", "age_geq_60", "Gender", "caucasian", "wscore_geq_3", "april", "may", "june", "july", "aug", "sep", "oct"))
```
The estimated VE and their 95\% confidence intervals are:

* Logistic regression: `r round(pfizer_nc$VE_logit, 3)` (`r round(pfizer_nc$VE_CI_logit[1], 3)`, `r round(pfizer_nc$VE_CI_logit[2], 3)`)
* IPTW: `r round(pfizer_iptw$ve, 3)` (`r round(pfizer_iptw$ve_ci[1], 3)`, `r round(pfizer_iptw$ve_ci[2], 3)`)
* NC estimator: `r round(pfizer_nc$VE, 3)` (`r round(pfizer_nc$VE_CI[1], 3)`, `r round(pfizer_nc$VE_CI[2], 3)`)


2. Moderna vaccine
```{r moderna_nc}
moderna_data <- tnd_data_2 %>% filter(Vaxname %in% c("Moderna", "none"))

moderna_iptw <- tnd_iptw(data = list(A = moderna_data$vaccinated,
                                    Y = moderna_data$COVID,
                                    X = moderna_data[, c("age_geq_18_le_60", "age_geq_60", "Gender", "caucasian", "wscore_geq_3", "april", "may", "june", "july", "aug", "sep", "oct")]))

moderna_nc <- nc_covid_ve(data = moderna_data,
                         nce = "IMM",
                         nco = "nco",
                         x = c("age_geq_18_le_60", "age_geq_60", "Gender", "caucasian", "wscore_geq_3", "april", "may", "june", "july", "aug", "sep", "oct"))
```
The estimated VE and their 95\% confidence intervals are:

* Logistic regression: `r round(moderna_nc$VE_logit, 3)` (`r round(moderna_nc$VE_CI_logit[1], 3)`, `r round(moderna_nc$VE_CI_logit[2], 3)`)
* IPTW: `r round(moderna_iptw$ve, 3)` (`r round(moderna_iptw$ve_ci[1], 3)`, `r round(moderna_iptw$ve_ci[2], 3)`)
* NC estimator: `r round(moderna_nc$VE, 3)` (`r round(moderna_nc$VE_CI[1], 3)`, `r round(moderna_nc$VE_CI[2], 3)`)


3. Johnson \& Jognson's Janssen vaccine
```{r janssen_nc}
janssen_data <- tnd_data_2 %>% filter(Vaxname %in% c("Janssen", "none"))

janssen_iptw <- tnd_iptw(data = list(A = janssen_data$vaccinated,
                                    Y = janssen_data$COVID,
                                    X = janssen_data[, c("age_geq_18_le_60", "age_geq_60", "Gender", "caucasian", "wscore_geq_3", "april", "may", "june", "july", "aug", "sep", "oct")]))

janssen_nc <- nc_covid_ve(data = janssen_data,
                         nce = "IMM",
                         nco = "nco",
                         x = c("age_geq_18_le_60", "age_geq_60", "Gender", "caucasian", "wscore_geq_3", "april", "may", "june", "july", "aug", "sep", "oct"))
```
The estimated VE and their 95\% confidence intervals are:

* Logistic regression: `r round(janssen_nc$VE_logit, 3)` (`r round(janssen_nc$VE_CI_logit[1], 3)`, `r round(janssen_nc$VE_CI_logit[2], 3)`)
* IPTW: `r round(janssen_iptw$ve, 3)` (`r round(janssen_iptw$ve_ci[1], 3)`, `r round(janssen_iptw$ve_ci[2], 3)`)
* NC estimator: `r round(janssen_nc$VE, 3)` (`r round(janssen_nc$VE_CI[1], 3)`, `r round(janssen_nc$VE_CI[2], 3)`)